---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Introduction

Thisassignment is a brief analysis of the steps performed by an individual over two months measured at 5 minute intervals.

## Loading and preprocessing the data

```{r setup, echo=TRUE,message=FALSE}
#Read the data & call libraries
if(!file.exists("activity.csv")){
unzip("activity.zip")  
}
basedata = read.csv("activity.csv")
library(dplyr)
library(ggplot2)

```


```{r cars, echo=TRUE}
# Data transformation, total steps by date
newbase = basedata %>%
  group_by(date)%>%
  summarise(stepsbydate = sum(steps))
```

## What is mean total number of steps taken per day?

The average daily steps is 10,765, the histogram shows a distribution close to normal.


```{r}
summary(newbase$stepsbydate)
```


```{r, message=FALSE, warning=FALSE}
ggplot(newbase, aes(stepsbydate))+
  geom_histogram(aes(y=..density..))+
  geom_density(col="red")+
  labs(title = "Histogram of daily steps")
```


## What is the average daily activity pattern?

The graph and the table show a very marked peak in the 5-minute interval that corresponds to the time from 1:55 p.m. to 2:00 p.m., in fact, the following intervals in the list are also close to this time. During this mentioned period, on average each day this individual performs 206 steps.

```{r}
MeanSteps = basedata  %>%
  filter(is.na(steps) == FALSE)%>%
  group_by(interval)%>%
  summarise(Mean = round(mean(steps),0))

ggplot(MeanSteps, aes(x=interval, y=Mean))+
  geom_line()+
  labs(title="Average for each 5-minute interval")

```





```{r}
knitr::kable(head(arrange(MeanSteps, desc(Mean)),n = 10), align = 'c',
             caption = "Top ten of steps in 5 minutes intervals")

```


## Imputing missing values

The _'missing values'_ will be replaced by the average values for the corresponding five-minute interval, this is achieved first by calculating said average, and then substitute each missing value in the base.

```{r}
# Calculate the average for each 5-minute interval

basedataNA = basedata %>%
  group_by(interval)%>%
  mutate(newsteps = mean(steps, na.rm=TRUE))

# This part replaces the NAs by the average of the corresponding five-minute interval.

for (i in 1:length(basedataNA$newsteps)) {
  if (is.na(basedataNA$steps[i])==TRUE) {
    basedataNA$steps[i] <- basedataNA$newsteps[i]
  }
  
}

```



### Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

In this case, the distribution is clearly more focused on the mean and in effect there is no difference between the mean and the median.

```{r, message=FALSE}

newbaseNA = basedataNA %>%
  group_by(date)%>%
  summarise(stepsbydate = sum(steps))

summary(newbaseNA$stepsbydate)

ggplot(newbaseNA, aes(stepsbydate))+
  geom_histogram(aes(y=..density..))+
  geom_density(col="red")

```

## Are there differences in activity patterns between weekdays and weekends?

The graph shows the average differences from weekends. During Saturday and Sunday the morning activity clearly decreases and on the other hand, in the daytime periods there is more activity on the weekends.

```{r}
basedata$date <- as.Date(basedata$date)
basedata$DoW <- weekdays(basedata$date)
basedata$WeekEnd = "No"
basedata$WeekEnd[basedata$DoW == "sábado" | basedata$DoW=="domingo"] = "Yes"

MeanStepsNA = basedata  %>%
  filter(is.na(steps) == FALSE)%>%
  group_by(interval, WeekEnd)%>%
  summarise(Mean = mean(steps))

ggplot(MeanStepsNA, aes(x=interval, y=Mean))+
  geom_line(aes(col=WeekEnd))

```